---
title: "Covid91 vaccine study Final2024"
author: "Addler Fuller"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males
```{r}
Males <-filter(FinalDat, sex=="M")
```


```{r}
Males <-select(Males,-sex)
```

## Graphic Desscription 
```{r}
dd2 <- Males %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

The male population is represented in this plot along with its relationship to the Covid 19 infection by using either a placebo or the vaccine. 

The data above data shows that 25% of vaccine recipients actually developed COVID-19, compared to only about 75% of placebo recipients.  Those who received the vaccine had a much lesser chance of contracting COVID-19 than those who received a placebo. It seems that the vaccine offers no defense against COVID-19. While receiving the Placebo, there is a 3:1 chance of contracting COVID-19. Men who used the medication had a 1.5-fold lower risk of contracting COVID-19 compared to those who took a Placebo.

Two plots are present. The x-axis in the first plot indicates infected, denoting whether or not the participants have COVID-19 infection. The count, or total number of individuals who received COVID-19 or not, is shown on the y-axis. In the above-mentioned second plot, The All of the individuals with COVID-19 are represented by the graph's left two bars. Whether or not a person has COVID-19 is shown on the x-axis. However, the y-axis now shows the percentage of patients who contracted COVID-19 after receiving the vaccine or a placebo, rather than the total number of cases of the disease.

There are two plotlines in this story. In the first plot, the x-axis shows infected, which indicates if the participants are infected with COVID-19. On the y-axis is the count, or total number of people who received COVID-19 or not. In the second plot described above, the graph's left two bars reflect each individual with COVID-19. The x-axis indicates a person's COVID-19 status. Instead of the overall number of cases of the illness, the y-axis now displays the proportion of patients who got COVID-19 after getting the vaccination or a placebo.

### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=Males)
rowPerc(table1)
colPerc(table1)
```


Based on the numerical data, 74% of males who took a placebo and 26% who received the vaccine contracted COVID-19. When receiving the vaccine, men are less likely to get COVID-19 than when receiving a placebo. The vaccination is a more effective medication to take than a placebo. The Placebo has a twice as high chance of contracting COVID-19.

### Inferential 

```{r}
chisq.test(table1)
chisqtestGC(table1)
Malesfish <- fisher.test(table1)
Malesfish
```


The output of the Fisher's Exact test should be carefully examined among all of these findings. At 2,2e-16, the p-value shown above is incredibly small. This is a remarkably little figure. This figure indicates that it is unlikely that these outcomes happened by accident.

According to the null hypothesis, sex has no bearing on the mortality caused by COVID-19. The alternative theory—that there is a relationship between sex and COVID-19 deaths—is the more likely one given the overwhelming evidence that sex does affect people's death rates.


```{r}
effect <- Malesfish$p.value/(Malesfish$p.value +1)
effectivenessMales = Malesfish$estimate/(Malesfish$estimate +1)*100
print(effectivenessMales)
```

```{r}
low=(Malesfish$conf.int[1]/(Malesfish$conf.int[1] +1))*100
hi=(Malesfish$conf.int[2]/(Malesfish$conf.int[2] +1))*100
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`

The efficacy of the COVID-19 vaccine for men is seen above. The vaccine is effective if the efficacy rate is higher than 50%. We can conclude that the vaccine is unsuccessful in preventing men from contracting COVID-19 based on the interval shown above. 

The approximate effectiveness of the vaccine is indicated by the odds ratio. As can be seen above, the men' odds ratio is 74. For the males, the effectiveness confidence interval is 78.32 to 68.84.



## Females

```{r}
Females <-filter(FinalDat, sex=="M")
```

```{r}
Females <-select(Females,-sex)
```

```{r}
dd2 <- Females %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```



## LGBTQ



## Druggies


# Overall Results and Conclusions